# Calculate Percent Cover of Hydrophytic Species

```{r include=FALSE}
library(tidyverse)
library(tidyr)
library(sf)
library(s2)
library(mapview)

mapviewOptions(fgb=F)

knitr::opts_chunk$set(echo = TRUE)
```

## Load Data

```{r }
lpi_CorrectCodes<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/lpi_tall_CorrectCodes.csv") %>% 
  select(LineKey,
         RecKey,
         FormType,
         FormDate,
         CheckboxLabel,
         PrimaryKey,
         PointLoc,
         PointNbr,
         layer,
         SpeciesCode,
         chckbox,
         source,
         SpeciesState,
         Scientific.Name.with.Author,
         CorrectSpeciesCode2)

WetAIMmasterList<- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/WetIndicators/WetlandAIM_MasterSpeciesList.csv") %>% 
  select(Symbol,
         WMVC_WetStatus,
         AW_WetStatus,
         GP_WetStatus)

Regions<-st_read("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/coe_regions/USACE_Regions_NAD83.shp")

header <- readRDS("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/AIM_tall_tables_export_2021-09-21/header.Rdata")

```

## Remove Soil Surface Codes and non plant codes

Remove all codes that are not vascular plants, so that I can calculate relative % cover

```{r}
lpi_CorrectCodesPlants <- lpi_CorrectCodes %>% 
  rename("CorrectSpeciesCode" = "CorrectSpeciesCode2") %>% 
  filter(layer != "SoilSurface",
         !SpeciesCode %in% c("L", "WL", "VL", "WA", "S"))

```


## Calculate the percent relative cover of each species for each transect

After checking for NAs in LineKey, it looks like every entry has a value for that column, so I will go with that.

```{r}
RelCover<-lpi_CorrectCodesPlants %>% 
  group_by(PrimaryKey, LineKey) %>% 
  mutate(NumPlantHits = length(LineKey)) %>% 
  group_by(PrimaryKey, LineKey,CorrectSpeciesCode) %>% 
  mutate(NumSpeciesHits = length(CorrectSpeciesCode)) %>% 
  summarise(RelativeCover = NumSpeciesHits/NumPlantHits) %>% 
  distinct(.)
 
#I think this worked, though there was likely a much easier way to do it
```

## Apply Wetland Indicator Statuses

```{r}
RelCoverWetIndicators<-RelCover %>% 
  left_join(WetAIMmasterList, by = c("CorrectSpeciesCode" = "Symbol"))
```

## Spatially apply the wetland indicator regions

### Find the LRRs of every Primary Key 

Use the lat long from the header files to match LRRs to each plot

```{r}
#create a spatial object from the LRR file
st_crs(Regions)
  
mapview(Regions)

#Create a spatial object using the coordinates from the header file
SpatialInfo<-header %>% 
  st_as_sf(., coords = c("Longitude_NAD83", "Latitude_NAD83"), crs = 4269) %>% 
  select("PrimaryKey")

#tell it to use s2 to get rid of the planar/projection issue
sf_use_s2()

#Join those two files
WetRegions<-st_join(Regions, SpatialInfo, join = st_contains)

#Convert the new file to a data frame to be merged with the lpi data
WetRegions<-as.data.frame(WetRegions) %>% 
  select(-geometry)

```

### Join the LRR data to the LPI data

```{r}
#Join the dfs and remove all the sites with Alaska LRRs to try to get rid of all Alaska sites
RelCovWetIndicatorswLRR<- RelCoverWetIndicators %>% 
  left_join(WetRegions)  %>% 
  filter(Region != "USACE Alaska Region") %>% 
  mutate(WetIndicator = case_when(
    Region == "USACE Arid West Region" ~ AW_WetStatus,
    Region == "USACE Western Mountains, Valleys, and Coast Region" ~ WMVC_WetStatus,
    Region == "USACE Great Plains Region" ~ GP_WetStatus)) %>% 
  select(-WMVC_WetStatus,
         -AW_WetStatus,
         -Region)

#It seems like using s2 eliminated the issue of some Regions being NA
```

## Calculate the percent hydrophytic per plot

```{r}

RelCovHydro<- RelCovWetIndicatorswLRR %>% 
  filter(WetIndicator == "FAC" | WetIndicator == "FACW" | WetIndicator == "OBL") %>% 
  group_by(PrimaryKey,LineKey)  %>% 
  summarise(TransectPercent = sum(RelativeCover)) %>% 
  ungroup()

RelCovHydroPlot<- RelCovHydro %>% 
  group_by(PrimaryKey) %>% 
  summarise(PlotPercent = mean(TransectPercent))

#Plots that have over 50% hydrophytic
over50 <- RelCovHydroPlot %>% 
  filter(PlotPercent > 0.5)

#Plots that have between 25 and 50% hydrophytic
between25and50 <- RelCovHydroPlot %>% 
  filter(PlotPercent > 0.25 & PlotPercent <= 0.5)

```

## Reattach the spatial info 

### Reattach the lat long info sourced from header
```{r}
locationinfo<-header %>% 
  select(PrimaryKey, Latitude_NAD83, Longitude_NAD83)

over50 <- over50 %>% 
  left_join(locationinfo)

between25and50 <- between25and50 %>% 
  left_join(locationinfo)

rm(list=ls())
```
