# Surface Water

```{include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Load the LPI data
```{r}
header <- readRDS("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/AIM_tall_tables_export_2021-09-21/header.Rdata") %>% 
  select(PrimaryKey, State)

lpi_tall<-readRDS("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/AIM_tall_tables_export_2021-09-21/lpi_tall.Rdata") %>% 
  select(LineKey, PrimaryKey, PointLoc, PointNbr, layer, code) %>% 
  left_join(header) %>% #attach state from header so you can remove alaska data
  filter(State != "AK")

```
## Calculate percent surface water

```{r}

SurfaceWater<-lpi_tall %>%
  group_by(PrimaryKey, LineKey) %>% 
  mutate(NumLPIHits = n_distinct(PointNbr)) %>% 
  mutate(WaterHits = if_else(code %in% c("W", "WA"), 1, 0)) %>% 
  group_by(PrimaryKey, LineKey) %>% 
  mutate(NumWaterHits = sum(WaterHits)) %>% 
  summarise(PercentWaterTransect = NumWaterHits/NumLPIHits) %>% 
  distinct(.) %>%  
  group_by(PrimaryKey) %>% 
  summarise(PercentWaterPlot = mean(PercentWaterTransect)) %>% 
  filter(PercentWaterPlot != 0)

over2<-SurfaceWater %>% 
  filter(PercentWaterPlot>0.02)

over1<-SurfaceWater %>% 
  filter(PercentWaterPlot>0.01)
```
I double checked a handful of the percentages and they look good! Ready to move on.

## Join this data with the data of 25 to 50% hydrophytic species

Since I used the original data to do this, I need to now make sure all of those sites also have 25-50% hydrophytic species.  Having skipped ahead (oops), it seems that there are definiteley a number that do not have 25-50%.  Do I want to throw out the plots that have less than that?  OH! Or, is there a chance some of them have greater than 50%?  I think I want to start out by just looking at the percentages to see what I am working with, so I will first join with all of the hydrophytic data somehow..

```{r}
RelHydroCover<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/RelHydroCover.csv")

over2hydroCover<-over2 %>% 
  left_join(RelHydroCover)

rm(list=ls())
```

