# Initial Analyses

```{include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(rgbif)
library(ggthemes)
```

## Load the data

### Wetland AIM data

```{r}
LPIcover2019_CO<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/WetlandAIM/LPIcover2019_CO.csv") %>% 
  mutate(year = 2019)

LPIcover2019_UT<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/WetlandAIM/LPIcover2019_UT.csv") %>% 
  mutate(year = 2019)

#these are all *absolute metrics*
WetAIMIndicators2020<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/WetlandAIM/WetAIMIndicators2020.csv") %>% 
  mutate(year = 2020) %>% 
  rename(Plot.ID = PlotID)

WetAIM.Relative.Cover2020<- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/WetlandAIM/ALL_RelativeCoverMetrics2020.csv") %>% 
  mutate(year = 2020) %>% 
  rename(Plot.ID = PlotID)

#why are there two hydrophyte columns??
hydrophytes<-WetAIMIndicators2020 %>% 
  select(Plot.ID, CommunityLPIPctHydroFAC, HydroFACCover)

```

### Terrestrial AIM data 

```{r}
#load this info for later use
header <- readRDS("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/AIM_tall_tables_export_2021-09-21/header.Rdata")
StateSpecies<- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/ExportedTerrestrial_Data/tblStateSpecies.csv")
RelativeCover<- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/RelativeCover.csv")
lpi_CorrectCodes<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/lpi_tall_CorrectCodes.csv")

#this is the table that contains elevation data for each transect
tblLines <- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/ExportedTerrestrial_data/tblLines.csv")

#load the final 25-50 hydrophytic data that matched the queries
all25to50<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/Btwn25and50hydrophytic/all25to50.csv")

over50<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/over50hydrophytic.csv") %>% 
  select(PrimaryKey, PlotPercent)

#Load all of the points with 25-50% (before running queries) to get the percent hydrophytic back to the all25to50 data
between25to50hydrophytic<-read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/25to50hydrophytic.csv") %>% 
  select(PrimaryKey, PlotPercent)

#left join to get the percent hydrophytic on to all25to50
all25to50<-all25to50 %>% 
  left_join(between25to50hydrophytic) 

#combine them to get a single dataset of "wetlands"

Terr.Wetlands<-rbind(all25to50, over50) %>% 
  mutate(source = "Terrestrial") %>% 
  rename(HydroCover = PlotPercent) %>% 
  rename(Plot.ID = PrimaryKey)
```

## Percent Hydrophytic 

### Terrestrial "wetlands" sites with Wetland AIM sites

```{r}
#pull out the hydrophytic cover from the wetland AIM datasets 
CO2019.hydroCover<-LPIcover2019_CO %>% 
  select(Plot.ID, Relative.HydrophyteFAC.Cover) %>% 
  rename(HydroFACCover = Relative.HydrophyteFAC.Cover)

UT2019.hydroCover<-LPIcover2019_UT%>% 
  select(Plot.ID, Relative.HydrophyteFAC.Cover) %>% 
  rename(HydroFACCover = Relative.HydrophyteFAC.Cover)

all2020.hydroCoverREL<-WetAIM.Relative.Cover2020 %>% 
  select(Plot.ID, RelativeHydroFACCover) %>% 
  rename(HydroFACCover = RelativeHydroFACCover) %>% 
  mutate(HydroFACCover = HydroFACCover/100)

#merge all of the Wetland data together
WetAIMhydrophytic<-rbind(CO2019.hydroCover, UT2019.hydroCover, all2020.hydroCoverREL) %>% 
  mutate(source = "Wetland") %>% 
  rename(HydroCover = HydroFACCover) 

#now combine all the data for simpler plotting
all.sites<-rbind(Terr.Wetlands, WetAIMhydrophytic)

ggplot()+
  geom_boxplot(data = all.sites, aes(x = source, y = (HydroCover*100)))+
  theme_bw()+
  labs(x = "AIM Program", 
       y = "Percent Hydrophytic Species", 
       title = "Relative Hydrophytic Cover")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))

```
### All Terrestrial sites with Wetland AIM sites

```{r}
#I think this only has sites that have FAC or wetter plants - go back to CalculatePercentCover.Rmd and change that somewhere...

TerrRelCovHydro <- read.csv("/Users/elinbinck/Documents/Grad_School/Thesis/R_project/Thesis_Research/data/RelHydroCover.csv") %>% 
  mutate(source = "Terrestrial") %>% 
  rename(HydroCover = PlotPercent, Plot.ID = PrimaryKey)

all.sites.allTerr<-rbind(TerrRelCovHydro, WetAIMhydrophytic)

ggplot()+
  geom_boxplot(data = all.sites.allTerr, aes(x = source, y = (HydroCover*100)))+
  theme_bw()+
  labs(x = "AIM Program", 
       y = "Percent Hydrophytic Species", 
       title = "Relative Hydrophytic Cover")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))

```

## Percent Native/Noxious

### Calculate Relative cover for Terrestrial wetlands

Here I will need to attach states from Header, and if species are noxious from state species, assuming these are state specific.  I will also need to load in the lpi tall table info with all of the species in order to calculate the metrics.  

```{r}
#Attach states to the Terrestrial data 
#first, extract the state for each plot
states<-header %>% 
  select(PrimaryKey, State)

#then, add that info to the terrestrial wetlands
Terr.Wetlands.state<-Terr.Wetlands %>% 
  left_join(states, by = c("Plot.ID" = "PrimaryKey"))

#next match up all the lpi data for those plots
Terr.Wet.lpi<-Terr.Wetlands.state %>% 
  inner_join(lpi_CorrectCodes, by = c("Plot.ID" = "PrimaryKey")) %>% 
  select(Plot.ID, HydroCover, source.x, State, LineKey, layer,SpeciesCode, CorrectSpeciesCode2, PointNbr)

#create an object with species codes, whether they are noxious or not, and the corresponding state for that info
noxious.info<- StateSpecies %>% 
  select(SpeciesCode, Noxious, SpeciesState) %>% 
  distinct(.)

unique(noxious.info$Noxious)

#try to join the noxious info to the lpi data... hopefully this won't create duplicates?
Terr.Wet.noxious<-Terr.Wet.lpi %>% 
  left_join(noxious.info, by =c("SpeciesCode", "State" = "SpeciesState"))

#Now calculate the number of noxious species and calculate relative percent cover
Terr.Rel.Noxious<-Terr.Wet.noxious %>% 
  group_by(Plot.ID, LineKey) %>% 
  mutate(NumPlantHits = length(LineKey)) %>% 
  mutate(Noxious = if_else(Noxious == "YES" | Noxious == "Yes", 1, 0, missing = 0)) %>% 
  group_by(Plot.ID, LineKey) %>% 
  summarise(NumNoxiousHits = sum(Noxious), NumPlantHits) %>% 
  summarise(RelativeNoxCoverLine = NumNoxiousHits/NumPlantHits) %>% 
  group_by(Plot.ID) %>% 
  summarise(RelativeNoxCoverPlot = mean(RelativeNoxCoverLine)) %>% 
  mutate(source = "Terrestrial")
  
```

To be continued...